Inferring ecological interactions

### 2016-12-07 SimulatedData.Rmd
link to manuscript

```{r,warning=FALSE,message=FALSE,echo=FALSE,cache=FALSE}
library(reshape2)
library(bipartite)
library(ggplot2)
library(knitr)
library(dplyr)
library(stringr)
library(R2jags)
opts_chunk$set(message=FALSE,warning=FALSE,fig.width=5,fig.height=5,echo=TRUE,cache=F,fig.align='center',fig.path="figure/")
```

@@

  #regression slopes
  #traits
  beta1_mu<- -1

@@

        #intensity
        N[x,y]<-exp(alpha[x] + beta1[x] * traitmatch[x,y])

        #draw one state
        obs[x,y]<-rpois(1,N[x,y])
    }
  }

@@

traitplot<-function(dat){
    ggplot(data=dat,aes(y=Interactions,x=abs(TraitI-TraitJ))) + geom_point() + geom_smooth(aes(group=1),method="glm",method.args=list(family="poisson"))
}

traitplot(dat)
```

# Poisson Regression

```{r,eval=T}
#format traitmatch as matrix
dat$Traitmatch<-abs(dat$TraitI-dat$TraitJ)
Traitmatch<-acast(data=dat,I~J,value.var="Traitmatch")

runs<-100000

#Source model
source("Bayesian/Poisson.R")

#print model
print.noquote(readLines("Bayesian//Poisson.R"))

@@

  #Parameters to track
  ParsStage <- c("beta1","ynew","fit","fitnew")

@@

    m1<-do.call(jags.parallel,list(data=modelDat,parameters.to.save=ParsStage,model.file="Bayesian/Poisson.jags",n.thin=nt, n.iter=ni,n.burnin=nb,n.chains=nc,DIC=F))
```

```{r}
parsO<-melt(m1$BUGSoutput$sims.array)
colnames(parsO)<-c("Draw","Chain","parameter","estimate")

#label species and plants
l<-levels(parsO$parameter)

#parameters to save
totrack<-m1$parameters.to.save

#assign species index to ragged frame.
sp_pl<-data.frame(parameter=l,Index=as.numeric(str_match(l,pattern="\\[(\\d+)]")[,2]),par=str_extract(l,"\\w+"))

#merge levels
pars<-merge(parsO,sp_pl)

pars<-pars[!pars$par %in% "deviance",]

@@

```{r,fig.width=15}
#view chaings
ggplot(pars[pars$par %in% c("beta1"),],aes(x=Draw,col=as.factor(Chain),y=estimate))  + geom_line() + facet_wrap(~Index,scales="free") 
```

##Posterior predictive check

```{r,fig.height=4,fig.width=8}
fitstat<-droplevels(pars[pars$par %in% c("fit","fitnew"),])
fitstat<-dcast(fitstat,Draw+Chain~par,value.var="estimate")

#add 1:1 line
ymin<-round(min(c(fitstat$fit,fitstat$fitnew)))
ymax<-round(max(c(fitstat$fit,fitstat$fitnew)))
ab<-data.frame(x=ymin:ymax,y=ymin:ymax)
p<-ggplot(fitstat,aes(x=fit,y=fitnew)) + geom_point() + theme_bw() + coord_equal()
psim4<-p  + labs(x="Discrepancy of observed data",y="Discrepancy of replicated data") + geom_line(data=ab,aes(x=x,y=y)) + ggtitle("Simulated Data")
psim4

#View simulated and predicted traitmatching

```{r}

#merge the dataframes.
dat$Index<-1:nrow(dat)

@@

# Compute emergent statistics

```{r}

makeN<-function(x){
  predweb<-acast(data=x,I~J,value.var="estimate")

  #calculate network statistic
  nstat<-networklevel(predweb,index=c("weighted connectance","weighted NODF"))
  data.frame(Metric=names(nstat),nstat)
}

@@

#calculate true network stat
tweb<-acast(dat,I~J,value.var="Interactions")
tstat<-networklevel(tweb,index=c("weighted connectance","weighted NODF"))
tstat<-data.frame(Metric=names(tstat),tstat)

ggplot(nstats,aes(x=nstat)) + geom_density(alpha=0.6,fill="black") + facet_wrap(~Metric,scales='free',nrow=2) + theme_bw()  + geom_vline(col="red",linetype="dashed",data=tstat,aes(xintercept=tstat)) + labs(x="value")
```

>>>>

```{r,warning=FALSE,message=FALSE,echo=FALSE,cache=FALSE}
library(reshape2)
library(bipartite)
library(ggplot2)
library(knitr)
library(boot)
library(tidyr)
library(dplyr)
library(stringr)
library(R2jags)
opts_chunk$set(message=FALSE,warning=FALSE,fig.width=5,fig.height=5,echo=TRUE,cache=F,fig.align='center',fig.path="figure/")

#source helper function
source('Bayesian/Functions.R')
```

@@

  #regression slope
  #traits
  beta1_mu<- -1

@@

        #intensity
        N[x,y]<-inv.logit(alpha[x] + beta1[x] * traitmatch[x,y])

        #draw one state
        obs[x,y]<-rbinom(1,1,N[x,y])
    }
  }

@@

traitplot<-function(dat){
    ggplot(data=dat,aes(y=Interactions,x=abs(TraitI-TraitJ))) + geom_point() + geom_smooth(aes(group=1),method="glm",method.args=list(family="binomial")) + labs(x="Difference in traits",y="Probability of Interaction")
}

traitplot(dat) + theme_bw()
```

# Intercept

```{r,eval=T}
fitb<-function(dat,algorithm="Intercept"){

  #format traitmatch as matrix
  dat$Traitmatch<-abs(dat$TraitI-dat$TraitJ)
  Traitmatch<-acast(data=dat,I~J,value.var="Traitmatch")

  if(algorithm == "Intercept"){

  runs<-10000

  #Source model
  source("Bayesian/Intercept.R")

  #print model
  print.noquote(readLines("Bayesian//Intercept.R"))

@@

    modelDat<-list("Yobs","Bird","Plant","Plants","Birds","Nobs")

    ParsStage <- c("alpha","alpha_mu","ynew","fit","fitnew")

m1<-do.call(jags.parallel,list(data=modelDat,parameters.to.save=ParsStage,model.file="Bayesian/Intercept.jags",n.thin=nt, n.iter=ni,n.burnin=nb,n.chains=nc,DIC=F))
} else

    if(algorithm=="Binomial") {

  runs<-10000

  #Source model
  source("Bayesian/Binomial.R")

  #print model
  print.noquote(readLines("Bayesian//Binomial.R"))

    #Parameters to track
    ParsStage <- c("alpha","beta","alpha_mu","beta_mu","ynew","fit","fitnew")

@@

m1<-do.call(jags.parallel,list(data=modelDat,parameters.to.save=ParsStage,model.file="Bayesian/Binomial.jags",n.thin=nt, n.iter=ni,n.burnin=nb,n.chains=nc,DIC=F))  
            }

    m1$Algorithm<-algorithm
    return(m1)
}
```

# Interecpt Model
```{r}
intercept<-fitb(dat,"Intercept")
```

# Binomial Regression
```{r}
binomial<-fitb(dat,"Binomial")
```

#Extract parameters from the model

```{r}
getpars<-function(m1,n.iter=100,nt=1,nb=0,pars=c("beta1","alpha1")){  
  parsm1<-extract_par(m1)
  parsm1$Model<-m1$algorithm
  return(parsm1)
}
```

```{r}
#bind them together
datpars<-list(intercept,binomial)

datpars<-lapply(datpars,getpars)
```

#Predict new species

```{r}
size_x = 2
size_y = 5

newtraits<-simA(size_x=size_x,size_y=size_y,traits_x=traits_x,traits_y=traits_y)

#calculate traitmatch
newtraits$Traitmatch<-abs(newtraits$TraitJ-newtraits$TraitI)

predictM<-function(m1,newtraits){
  parsm1<-extract_par(m1)

  #matching function
a<-parsm1 %>% filter(par %in% c("alpha_mu","beta_mu")) %>% select(Draw,Chain,par,estimate) %>% dcast(.,Draw + Chain ~ par,value.var="estimate") %>% group_by(Draw,Chain)%>%  do((pred.binomial(alpha=.$alpha_mu,beta=.$beta_mu,newdata=newtraits))) %>% group_by(I,J) %>% summarize(mean=mean(value),lower=quantile(value,0.95),higher=quantile(value,0.05))

return(a)
}

predt<-predictM(binomial,newtraits)
```

# Model evaluation

```{r}
parsall<-rbind_all(list(parsm1,parsm2))
```

## Posterior predictive check

```{r,fig.height=4,fig.width=8}
fitstat<-droplevels(parsall[parsall$par %in% c("fit","fitnew"),])
fitstat<-dcast(fitstat,Draw+Model+Chain~par,value.var="estimate")

#add 1:1 line
ymin<-round(min(c(fitstat$fit,fitstat$fitnew)))
ymax<-round(max(c(fitstat$fit,fitstat$fitnew)))
ab<-data.frame(x=ymin:ymax,y=ymin:ymax)

ggplot(data=fitstat,aes(x=fit,y=fitnew)) + geom_point(aes(col=Model)) + theme_bw() + coord_equal() + labs(x="Discrepancy of observed data",y="Discrepancy of replicated data") + geom_line(data=ab,aes(x=x,y=y)) + ggtitle("Simulated Data")
```

## Proportion of predicted links

```{r}
propc<-function(x,orig=dat){
  predweb<-acast(data=x,I~J,value.var="Index")

  #calculate network statistic
  nstat<-networklevel(predweb,index=c("connectance","nestedness"))
  data.frame(Metric=names(nstat),nstat)
}

prop_links<-parsall %>% filter(par=="ynew") %>% group_by(Model,Draw,Chain) %>% inner_join(dat) %>% do(propc(.))

@@

###View trait-matching

```{r}
#merge the dataframes.
dat$Index<-1:nrow(dat)

@@

### Compute emergent statistics

```{r}
makeN<-function(x){
  predweb<-acast(data=x,I~J,value.var="estimate")
  #calculate network statistic
  nstat<-networklevel(predweb,index=c("connectance","nestedness"))
  data.frame(Metric=names(nstat),nstat)
}

@@

#calculate true network stat
tweb<-acast(dat,I~J,value.var="Interactions")
tstat<-networklevel(tweb,index=c("connectance","nestedness"))
tstat<-data.frame(Metric=names(tstat),tstat)

ggplot(nstats[nstat$Metric %in% "connectance",],aes(x=nstat)) + geom_density(alpha=0.6,fill="black") + facet_wrap(~Metric,scales='free',nrow=1) + theme_bw()  + labs(x="value")
+ geom_vline(col="red",linetype="dashed",data=tstat,aes(xintercept=tstat)) 
```

### 2016-12-07 ms/intro_drafted.Rmd
link to manuscript

Biotic interactions form the backbone of biodiversity (Bascompte & Jordano 2007). They pose selective pressures that are key to evolution (ref) and operate together with abiotic factors in constraining the distribution of biodiversity (Soberon et al. 2007). Documenting and analyzing ecological networks is essential to: (i) understanding the rules underlying pairwise species interactions (Bascompte et al. 2006), (ii) quantifying ecosystem functioning and services (Thomson et al 2012), or (iii) predicting changes in from of global change scenarios (ref). In spite of increasing efforts, it is unlikely that we will ever manage to empirically document the massive amount of interactions existing in nature (Morales-Castilla et al. 2015). Because of this, complementing field work with modelling is gaining acceptance among network ecologists (ref). In a context where novel ecosystems arise as a result of species invasions or due to geographic shifts tracking climate change, our ability to predict future interactions will limit how efficiently we anticipate and respond to threats on biodiversity (i.e. species local extinctions) and on humans (i.e. spread of infectious disease or plagues affecting to crops).

>>>>

Biotic interactions form the backbone of biodiversity (Bascompte & Jordano 2007). They pose selective pressures that are key to evolution (ref) and operate together with abiotic factors in constraining the distribution of biodiversity (Soberon et al. 2007). Documenting and analyzing ecological networks is essential to: (i) understanding the rules underlying pairwise species interactions (Bascompte et al. 2006), (ii) quantifying ecosystem functioning and services (Thomson et al 2012), or (iii) predicting changes in from of global change scenarios (ref). In spite of increasing efforts, it is unlikely that we will ever manage to empirically document the enormous amount of interactions existing in nature (Morales-Castilla et al. 2015). Because of this, complementing field work with modelling of species interactions is gaining acceptance among network ecologists (ref). In a context where novel ecosystems arise as a result of species invasions or due to geographic shifts tracking climate change, our ability to predict future interactions will limit how efficiently we anticipate and respond to threats on biodiversity (i.e. species local extinctions) and on humans (i.e. spread of infectious disease or plagues affecting to crops).
